Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-11 popl24/automation.tex
checkpoint

\subsection{Brute Force Computation of Program Distributions}
\label{section-bruteforce}

We first explore a brute force approach to automatically verifying
program behaviour, the first step of which is to compute the basic
distribution of a program $\prog$. The program distribution
$\progd(\prog)$ can obviously be automatically derived from that, as
can be any of its marginal or conditional distributions. In this
Section, we consider two automated mechanisms for deriving
$\progd(\prog)$ in this way. The first is through a straightforward
computation of all final program memories.  The second approach
rewrites $\minifed$ programs to stratified Datalog programs, that are
amenable HPC optimizations such as parallelization and GPU matrix
computations as shown in recent work
\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.

@@

%The basic distribution of a program $\prog$ is based on the
%final memories resulting from any run of $\prog$, denoted
%$\runs(\prog)$.
%\begin{definition}
%  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
%  $$
%  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
%  $$
%\end{definition}

As we've observed, $\runs(\prog)$ can be represented in a truth table
form, as illustrated in Figure \ref{fig-basic-distributions}.

@@

Here we describe a straightforward technique for obtaining program
distributions $\progd(\prog)$ through direct computation of
$\runs(\prog)$. Although we could obtain $\runs(\prog)$ simply by
executing $\prog$ given all random tapes and input secrets, the method
we describe here is empirically more efficient (though still
exponential in the length of the random tape), and also supports the
translation to Datalog described in Section
\ref{section-bruteforce-datalog}. An implementation of these
techniques along with Examples from Sections
\ref{section-minicat-examples} and \ref{section-metalang-ygc} is
available online\footnote{URL redacted for double-blind review.}.

In Figure \ref{fig-solve} we define the algorithm $\solve{\stores}{\be}$
which filters $\stores$ based on satisfaction of $\be$. This algorithm
is fundamental to our brute force methods. It's correctness is characterized
as follows. 
\begin{lemma}
  \label{lemma-solves}
  For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
  $(\solve{\stores}{\be}) = \{ \store \in \stores \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
  for some $\cid$.
\end{lemma}
\begin{proof}
  Since we assume safety of programs we can assume that all variables in $\be$ have the
  same owner $\cid$. The result otherwise follows in a straightforward manner by induction
  on $\be$. 
\end{proof}

In order to automatically obtain $\runs(\prog)$, we can do a left
folding of $\solve$ across $\prog$, inductively extending memories
with valid view assignments in the order of their definition.
We can then use this to compute $\progd(\prog)$. We denote
these computations as $\cruns$ and $\cprogd$ respectively.
\begin{lemma}
  \label{lemma-cruns}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
  \begin{eqnarray*}
    {tt}\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) &\defeq& \begin{array}{l}
      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
  \end{eqnarray*}
  Then $\cruns(\prog) = \runs(\prog)$.
\end{lemma}
\begin{proof}
  By Lemma \ref{lemma-solves} and induction on the length of $\prog$. 
\end{proof}
\begin{lemma}
  \label{lemma-cprogd}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define
  $\cprogd(\prog)$ as the distribution $\pmf$
  where for all $\store \in \mems(S\cup F \cup V)$:
  $$
  \pmf(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
    \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
  $$
  Then:
  $$
  \cprogd(\prog) = \progd(\prog)
  $$
\end{lemma}
\begin{proof}
  Immediately by Lemma \ref{lemma-cruns} and Definition \ref{def-progd}.
\end{proof}

\subsection{Protocol Certification}

@@

We base certification of protocols $\prog$ on a calculation of its
program distribution. Recall from Definition \ref{definition-progd} that
program distributions are calculated from runs of a protocol given
any random tape and input secrets. That is:
\begin{definition}
  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$, define:
  $$
  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
  $$
\end{definition}

Given any $\prog$, the set $\runs(\prog)$ can be computed via brute
force methods. One is to fold a standard interpretation of logical
formulas across $\prog$, obtaining a truth table representation of
$\runs(\prog)$. We can also interpret $\prog$ as a stratified Datalog
program, with $\runs(\prog)$ being the set of least Herbrand models
obtained by representing different random tapes and input secrets as
different fact bases. This representation is amenable HPC
optimizations such as parallelization and GPU matrix computations as
shown in recent work
\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.
We provide more details of brute force computation of $\runs(\prog)$ in
Appendix \ref{section-bruteforce} and an implementation of
the tabular method is available in our online repository\footnote{Link
redacted for double-blind reviewing.}.

Given a calculation of $\runs(\prog)$, we can easily calculate its
basic distribution as in Definition \ref{definition-progd)}, as well
as any marginalization (such as $\progd(\prog)$) or conditioning of it.

@@

We can thus automatically verify security properties of examples considered
in previous sections, observing that passive security of $\prog_{\ref{example-OT}}$
has not been previously demonstrated.
\begin{lemma}
  $\prog_{\ref{example-he}}$, $\prog_{\ref{example-gmw-andcircuit}}$,
  $\prog_{\ref{example-ygc-andcircuit}}$ are all passive secure,
  $\prog_{\ref{example-OT}}$ is passive secure assuming client 3 is
  honest, and $\prog_{\ref{example-otp}}$ satisfies probabilistic
  noninterference and hence perfect secrecy as defined in
  \ref{barthe2019probabilistic}.
\end{lemma}
\begin{proof}
  We have certified $\nimo(prog_{\ref{example-he}})$, $\nimo(\prog_{\ref{example-gmw-andcircuit}})$,
  $\nimo(\prog_{\ref{example-ygc-andcircuit}})$, and $\nimo(prog_{\ref{example-OT}})$ assuming
  $3 \in H$, and $\pni(\prog_{\ref{example-otp}})$. The result follows by Lemmas
  \ref{lemma-pni} and \ref{lemma-nimo}.
\end{proof}

@@

Our approach is thus to design gate certifications that demonstrably
guarantee security for circuits built from them. Intuitively, we can
consider any gate as a mini-program and certify intensional properties
of its isolated program distribution. These intensional properties are
guaranteed secure in circuit compositions, as we demonstrate using
methods based on probabilistic separation logic
\ref{barthe2019probabilistic}. While this last step is manual, our
gate certifications are defined generally and can be used to certify
new gate implementations, and verification applies to a circuit of any
size. In Section \ref{section-composition} we will demonstrate our
approach using YGC and GMW as examples.

When we consider intensional properties
of gates in isolation, we will explicitly consider flip conditions in
distributions. Thus we will be directly concerned with basic distributions,
which we denote $\progtt(\prog)$, and we will often focus on views
defined in subprotocols. 
\begin{definition}
  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
  $\prog$ as defined in Definition \ref{def-progd}. For any program $\prog =
  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
  $\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
\end{definition}
Notions of probabilistic independence, aka \emph{separation}, as well
as correlation, are important to certify. We borrow the symbols $*$
and $\sim$ from \cite{barthe2019probabilistic} with the same
denotations.
\begin{definition}
  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\sep{\pmf}{X}{Y}$ iff for all
    $\store \in \mems(X \cup Y)$ we have
  $\margd{\pmf}{X \cup Y}(\store) =
  \pmf(\store_X) * \pmf(\store_Y)$
\end{definition}
We immediately observe that separation and correlation are certifiable,
since they are based on marginalization and conditioning of program
distributions. 
\begin{lemma}
  Both of the following are true:
  \begin{enumerate}
  \item $\sep{\progtt(\prog)}{X}{Y}$ is certifiable for any $X$ and $Y$.
  \item $\vc{\progtt(\prog)}{x}{y}$ is certifiable for any $x$ and $y$.
  \end{enumerate}
\end{lemma}

@@

Since we certify properties of gates in isolation with witness inputs,
our metatheory is concerned with preservation of important invariants
when certified components are ``spliced'' into larger program
contexts.  One important reasoning principle is related to
noninterference properties- specifically, if we show that gate views
enjoy a noninterference property wrt their free (input) variables,
probabilistic independence propagates to other variables in the
preceding program.
\begin{restatable}[Noninterference]{lemma}{noninterference}
  \label{lemma-noninterference}
  Given $\prog_1;\prog_2$ and $X = \iov(\prog_2) - \vdefs(\prog_2)$ and
  $Y \subseteq \vdefs(\prog_2)$. If $\sep{\progd(\prog_1;\prog_2)}{X}{Y}$
  then $\sep{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}{Y}$.
\end{restatable}

>>>>

%We first explore a brute force approach to automatically verifying
%program behaviour, the first step of which is to compute the basic
%distribution of a program $\prog$. The program distribution
%$\progd(\prog)$ can obviously be automatically derived from that, as
%can be any of its marginal or conditional distributions. In this
%Section, we consider two automated mechanisms for deriving
%$\progd(\prog)$ in this way. The first is through a straightforward
%computation of all final program memories.  The second approach
%rewrites $\minifed$ programs to stratified Datalog programs, that are
%amenable HPC optimizations such as parallelization and GPU matrix
%computations as shown in recent work
%\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.

%Here we describe a straightforward technique for obtaining program
%distributions $\progd(\prog)$ through direct computation of
%$\runs(\prog)$. Although we could obtain $\runs(\prog)$ simply by
%executing $\prog$ given all random tapes and input secrets, the method
%we describe here is empirically more efficient (though still
%exponential in the length of the random tape), and also supports the
%translation to Datalog described in Section
%\ref{section-bruteforce-datalog}. An implementation of these
%techniques along with Examples from Sections
%\ref{section-minicat-examples} and \ref{section-metalang-ygc} is
%available online\footnote{URL redacted for double-blind review.}.

Our first step in automating security analysis is to show how to
compute the basic program distribution. Recall that basic
distributions are based on the memories resulting from execution of an
$\minicat$ program, which contain a record of all client views and the
public output.  Given $\prog$ we denote this as $\runs(\prog)$, and we
denote the basic distribution of a program as $\progtt(\prog)$.
\begin{definition}
  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
  $$
  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
  $$
  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
  $\prog$, where given $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$
  we have for all $\store \in \mems(S \cup V \cup F)$:
  $$
  \progtt(\prog)(\store) =  1 / 2^{|S\cup F|} \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
  $$
\end{definition}

Any marginalization or conditioning of $\progtt(\prog)$ can be easily
obtained algorithmically, in particular $\progd(\prog) =
\margd{\progtt(\prog)}{\iov(\prog)}$.  As already discussed in Section
\ref{section-hyperprop} and as we will discuss more below, extensional
and intensional security properties are predicated on marginalizations
and conditionings of basic distributions. So calculating
$\runs(\prog)$ establishes the foundation for our analysis.

@@

Intuitively, we can represent $\runs(\prog)$ in tabular form, as
illustrated in Figure \ref{fig-basic-distributions} for Examples
\ref{example-otp} and \ref{example-lambda-obliv}(a). So a basic
strategy is to directly compute this truth table for any given
program. We first describe an approach where we compute
$\runs(\instr_1,\ldots,\instr_n)$ iteratively for the subprograms
$\instr_1,\ldots,\instr_i$ for $1 \le i \le n$, by individually
computing the truth table of expressions $\be_i$ for each $\instr_i =
(\eassign{v_i}{\be_i})$, and joining this with
$\runs(\instr_1,\ldots,\instr_{i-1})$.

In Figure \ref{fig-solve} we define the algorithm
$\solve{\stores}{\be}$ which filters a given $\stores$ to obtain the
subset whose elements satisfy $\be$ (make it true). It's correctness
is characterized as follows.
\begin{lemma}
  \label{lemma-solves}
  For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
  $(\solve{\stores}{\be}) = \{ \store \in \stores \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
  for some $\cid$.
\end{lemma}
\begin{proof}
  Since we assume safety of programs we can assume that all variables in $\be$ have the
  same owner $\cid$. The result otherwise follows in a straightforward manner by induction
  on $\be$. 
\end{proof}

@@

To obtain $\runs(\prog)$, we can perform a left folding of $\solve$
across $\prog$ and each view definition, inductively extending
memories with valid view assignments in the order of their definition.
We denote this computation as $\cruns(\prog)$.
\begin{lemma}
  \label{lemma-cruns}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
  \begin{eqnarray*}
    {tt}\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) &\defeq& \begin{array}{l}
      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
  \end{eqnarray*}
  Then $\cruns(\prog) = \runs(\prog)$.
\end{lemma}
\begin{proof}
  By Lemma \ref{lemma-solves} and induction on the length of $\prog$. 
\end{proof}
The above yields an algorithm for computing basic distributions and
thus program distributions. 
\begin{lemma}
  \label{lemma-cprogd}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define
  $\cprogtt(\prog)$ as the distribution $\pmf$
  where for all $\store \in \mems(S\cup F \cup V)$:
  $$
  \pmf(\store) = 1 / 2^{|S\cup F|} \ \text{if}\ \store \in \cruns(\prog), \text{otherwise}\ 0
  %\begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
  %  \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
  $$
  Then $\cprogtt(\prog) = \progtt(\prog)$ and $\margd{\cprogtt(\prog)}{S\cup V} = \progd(\prog)$.
\end{lemma}
\begin{proof}
  Immediately by Lemma \ref{lemma-cruns} and Definition \ref{def-progd}.
\end{proof}
A full implementation of these methods is available in our online
repository\footnote{Link redacted for double-blind reviewing.}. To
avoid notational confusion we will continue to refer to $\progtt$ and
$\cruns$ in our subsequent discussion of automated verification but
with the understanding that both can be computed using the above
algorithms.

Since $|\runs(\prog)|$ grows exponentially in the size of
$\flips(\prog)$, scalability to larger programs is a concern.  However
computation of $\runs$ is amenable to optimizations, especially
parallelization- rather than computing the truth table
``horizontally'', it can be computed vertically. We can rewrite
$\minifed$ programs $\prog$ to stratified Datalog programs, where
$\runs(\prog)$ is equivalent to set of least Herbrand models obtained
by representing different random tapes and input secrets as different
fact bases. Details of this rewrite are provided in Appendix
\ref{section-bruteforce}. The model for each fact base can be computed
independently and in parallel, with the potential to apply other
matrix optimizations developed in recent work on HPC for Datalog
\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.

@@

@@

We can thus automatically verify security properties of examples considered
in previous sections, observing that passive security of $\prog_{\ref{example-OT}}$
has not been previously demonstrated.
\begin{lemma}
  $\prog_{\ref{example-he}}$, $\prog_{\ref{example-gmw-andcircuit}}$,
  $\prog_{\ref{example-ygc-andcircuit}}$ are all passive secure,
  $\prog_{\ref{example-OT}}$ is passive secure assuming client 3 is
  honest, and $\prog_{\ref{example-otp}}$ satisfies probabilistic
  noninterference and hence perfect secrecy (as defined in
  \cite{barthe2019probabilistic}).
\end{lemma}
\begin{proof}
  We have certified $\nimo(prog_{\ref{example-he}})$, $\nimo(\prog_{\ref{example-gmw-andcircuit}})$,
  $\nimo(\prog_{\ref{example-ygc-andcircuit}})$, and $\nimo(prog_{\ref{example-OT}})$ assuming
  $3 \in H$, and $\pni(\prog_{\ref{example-otp}})$. The result follows by Lemmas
  \ref{lemma-pni} and \ref{lemma-nimo}.
\end{proof}

@@

Our solution is to design gate certifications that demonstrably
guarantee security for circuits built from them. Intuitively, we can
consider any gate as a mini-program and certify intensional properties
of its isolated program distribution. These intensional properties are
guaranteed secure in circuit compositions, as we demonstrate using
methods based on probabilistic separation logic
\ref{barthe2019probabilistic}. While this last step is manual, our
gate certifications are defined generally and can be used to certify
new gate implementations, and verification applies to a circuit of any
size. In Section \ref{section-composition} we will demonstrate our
approach using YGC and GMW as examples.

When we consider intensional properties
of gates in isolation, we will explicitly consider flip conditions in
distributions. Thus we will be directly concerned with basic distributions,
which we denote $\progtt(\prog)$, and we will often focus on views
defined in subprotocols. 
Notions of probabilistic independence, aka \emph{separation}, as well
as correlation, are important to certify. We borrow the symbols $*$
and $\sim$ from \cite{barthe2019probabilistic} with the same
denotations.
\begin{definition}
  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\sep{\pmf}{X}{Y}$ iff for all
    $\store \in \mems(X \cup Y)$ we have
  $\margd{\pmf}{X \cup Y}(\store) =
  \pmf(\store_X) * \pmf(\store_Y)$
\end{definition}
We immediately observe that separation and correlation are certifiable,
since they are based on marginalization and conditioning of program
distributions. 
\begin{lemma}
  Both of the following are true:
  \begin{enumerate}
  \item $\sep{\progtt(\prog)}{X}{Y}$ is certifiable for any $X$ and $Y$.
  \item $\vc{\progtt(\prog)}{x}{y}$ is certifiable for any $x$ and $y$.
  \end{enumerate}
\end{lemma}

@@

Since we certify properties of gates in isolation with witness inputs,
our metatheory is concerned with preservation of important invariants
when certified components are ``spliced'' into larger program
contexts. This technique often requires us to focus on views defined
in subprograms, so we define the following.
\begin{definition}
For any program $\prog =
  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
$\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
\end{definition}

An important reasoning principle is related to component
noninterference- specifically, if we show that gate views enjoy a
noninterference property wrt their free (input) variables,
probabilistic independence propagates to other variables in the
preceding program.
\begin{restatable}[Noninterference]{lemma}{noninterference}
  \label{lemma-noninterference}
  Given $\prog_1;\prog_2$ and $X = \iov(\prog_2) - \vdefs(\prog_2)$ and
  $Y \subseteq \vdefs(\prog_2)$. If $\sep{\progd(\prog_1;\prog_2)}{X}{Y}$
  then $\sep{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}{Y}$.
\end{restatable}
